VISUALIZATION OF THE LEFT-SIDED RIEMANN-LIOUVILLE FRACTIONAL INTEGRAL \\ [3pt] IN ``FCAA'' JOURNAL

### 2019-10-30 fractional/FCAA/Authors Kit (Updated)/FCAA-template-for-authors.tex
worked in suggestions from CMJ

\noindent
In this article we specifically focus on the geometric interpretation of fractional integrals. To be more specific, we present a novel geometric interpretation of the left-sided Riemann-Liouville fractional integral $_{a'}I_t^{\alpha} f(t)$ based at $a'$ (see Section~\ref{sec:ls}) \cite{laurent1884}. We accomplish this by extending the geometric framework proposed by Podlubny \cite{podlubny02}.
%As we mentioned earlier, Riemann and Liouville played an important role in the development of fractional calculus. Their work was crucial in defining what is now known as 
%the Riemann-Liouville fractional integral \cite{laurent1884}. At this point in time the reader should be made aware of the fact that there exists many other accepted fractional integral definitions in the literature.
%We, however, deem these definitions to be beyond the scope of the current article and as such we do not discuss them any further in this work. 
%The geometric interpretation that we present in this paper here is an extension of the geometric interpretation presented by Podlubny  
Podlubny showed that a Riemann-Liouville fractional integral can be converted into a Riemann-Stieltjes integral. The Riemann-Stieltjes integral so obtained can then be visualized using Bullock's geometric framework \cite{bullock88}. Moreover, Grobler showed that a Riemann-Stieltjes integral can be interpreted as the area which is obtained by summing together the area of an infinite number of non-rectangular infinitesimals, i.e. a Riemann-Stieltjes integral can be converted into a Cavalieri integral \cite{ackermann12,grobler19}. Combining these two ideas results in the following geometric interpretation: the Riemann-Liouville fractional integral $_{a'}I_t^{\alpha} f(t)$ can be interpreted as the area which is obtained by summing together the area of an infinite number of non-rectangular infinitesimals. The shape
of the aforementioned infinitesimals is determined by the order of integration $\alpha$ and the integration limit $t$ of the fractional integral under consideration.\\

@@

\section{Riemann-Stieltjes integral}
\label{sec:rs}
\noindent
Let $f$ and $g$ denote real-valued functions defined on a closed interval $[a',b']$ of the real line. We shall suppose that both $f$ and $g$ are bounded on $[a',b']$. Recall that the Riemann-Stieltjes integral is defined as follows:
\begin{equation}
\label{eq:g_int2}
\int_{a'}^{b'} f(x) dg(x) =  \lim_{n \rightarrow \infty}\sum_{i=0}^{n-1} f(x_i)[g(x_{i+1})-g(x_{i})], 
\end{equation}
with $(x_i)_{i=0}^n$ being arbitrary partition points on the $x$-axis (such that $a' = x_0 \leq x_1 \leq \cdots \leq x_n = b'$). The function $g$ is known as the integrator and if it is monotone increasing then it maps points in the interval $[a',b,]$ to the interval $[a,b]$, with $a = g(a')$ and $b = g(b')$. The integral in equation~\eqref{eq:g_int2} is hard to evaluate in its present form. If $g$ is differentiable, however, then it becomes trivial to 
evaluate a Riemann-Stieltjes integral due to the following identity:
\begin{equation}
\label{eq:rs_identity}
\int_{a'}^{b'} f(x) dg(x) = \int_{a'}^{b'} f(x)g'(x)~dx.
\end{equation}

@@

\noindent
The definition of the Cavalieri integral is presented in Definition~\ref{def:cav_integral}. It is, however, our experience that Definition~\ref{def:cav_integral} is hard to assimilate if it is encounterd by a reader for the first time. We, therefore, encourage readers who are not familiar with the Cavalieri integral to rather first read Section~\ref{sec:cav_integral_example} instead. The reader who is not familiar with the Cavalieri integral should, however, feel free to refer back to the definitions presented in this section when things in Section~\ref{sec:cav_integral_example} become unclear. Moreover, the definitions presented below were taken from \cite{ackermann12}. Furthermore, none of the theorems in this section are proved, the proofs can be found in \cite{ackermann12,grobler19}.

@@

\noindent
Let $a(y)$ be a translational function with respect to a function $f(x)$ on the interval $[a,b]$. Moreover, $b(y) = a(y) + (b-a)$. Furthermore, the functions $a(y)$ and $b(y)$ intersect $f(x)$ at $a'$ and $b'$ respectively. These assertions are assumed to be true throughout the paper and will not be repeated again.

\begin{definition}\label{def:h}
The mapping $h : [a, b] \rightarrow [a',b']$, which maps $x_i^1 \in [a, b]$ to $x_i^2 \in [a',b']$, is defined as
$h(x_i^1) =$ $\{x_i^2 \in [a' ,b'] | a\circ f(x_i^2) + [x_i^1 - a] = x_i^2$ , $a = a(0)\}$
\end{definition}

\begin{definition}\label{def:g}
The mapping $g:[a', b'] \rightarrow [a, b]$, which maps $x_i^2 \in [a' , b']$ to $x_i^1\in [a, b]$,
is defined as $g(x_i^2) = x_i^2 - a \circ f (x_i^2) + a$.
\end{definition}

@@

\noindent
The functions $h(x)$ and $g(x)$ are known as the transformation function and the inverse transformation function, respectively. Furthermore,
\begin{equation}
\label{eq:g_def}
g(x) = x - a\circ f(x) + a 
\end{equation}
and
\begin{equation}
\label{eq:h_def}
h(x) = g^{-1}(x). 
\end{equation}

\begin{definition}\label{def:cav_integral}
Define a partition $\mathcal{P}_1$ of $[a,b]$ to be a set of points $x_0^1, x_1^1,\cdots,x_n^1$, 
where $a = x_0^1 \leq x_1^1 \leq \cdots \leq x_n^1 = b$. For each partition 
$\mathcal{P}_1$ of $[a,b]$ write $\Delta x_k^1 = x_k^1-x_{k-1}^1$. 
%Since both boundaries of any integration strip are necessarily
%translations of the translational function $a(y)$, we can apply the transformation
%function $h$ to the partition $\mathcal{P}_1$. 
If the transformation function $h$ is strictly increasing,
the application of $h$ to the partition $\mathcal{P}_1$ induces a new partition $\mathcal{P}_2 = \{x_0^2, x_1^2,\cdots, x_n^2\}$.
Otherwise, if $h$ is strictly decreasing, the application of $h$ induces a reversed partition $\mathcal{P}_2 = \{x_n^2, \cdots, x_1^2,x_0^2\}$. It
can be assumed that $h$ is strictly increasing, without any loss of generality.  Let 
$M_k = \sup \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, $m_k = \inf \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, and set 
\begin{equation}
\label{eq:c_up}
\mathcal{C}_U(\mathcal{P}_1,f,h) = \sum_{k=1}^n M_k \Delta x_k^1, 
\end{equation}
and
\begin{equation}
\label{eq:c_low}
\mathcal{C}_L(\mathcal{P}_1,f,h) = \sum_{k=1}^n m_k \Delta x_k^1. 
\end{equation}
The sums in equation~\eqref{eq:c_up} and equation~\eqref{eq:c_low} are respectively called the the upper and lower Cavalieri sums.
If there is a unique number $I$ that satisfies the inequality $\mathcal{C}_L(\mathcal{P}_1,f,h)\leq I \leq \mathcal{C}_U(\mathcal{P}_1,f,h)$ for all 
partitions $\mathcal{P}_1$ of $[a,b]$, then $I$ is called the Cavalieri integral of $f$ from $a(y)$ to $b(y)$ and is denoted by
\begin{equation}
\int_{a(y)}^{b(y)} f(x) dx.
\end{equation}
\end{definition}

@@

\noindent
Note that for all $K\in\mathbb{R}$ (see equation~\eqref{eq:rs_identity}):
\begin{equation}
\int_{a'}^{b'} f(x)\,d[\widetilde{g}(x)+K] = \int_{a'}^{b'}f(x)\,\widetilde{g}(x). 
\end{equation}
Furthermore, recall that
\begin{equation}
\label{eq:g_definition}
g(x) = x - a\circ f(x) + a.
\end{equation}
Moreover, let $C$ be a unique real-valued constant that satisfies: $g(x) := \widetilde{g}(x)+C$.\\

@@

\section{Cavalieri Integral: Clarifying Example}
\label{sec:cav_integral_example}
\noindent
The example which we present below was taken from \cite{ackermann12,grobler19}. Let the region $R$ be bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$ and $b(y)=4-y$. This region is depicted in Figure~\ref{fig:caval2}. Note that in the case of $R$: $a=1$, $b=4$, $a'=\frac{1}{2}$ and $b'=2$.\\

@@

\noindent
Let us express this idea more formally. Let $(x_i^1)_{i=0}^{n}$ denote a partition on the $x$-axis, such that $a = x_0^1 < x_1^1 < \cdots < x_n^1 = b$, and $\Delta x_i^1 = x_{i+1}^1 - x_i^1$. Consider the following lower Cavalieri sum (see equation~\ref{eq:c_low}):
\begin{equation}
\label{eq:cav_sum}
\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
The partition points $(x_i^1)_{i=0}^{n}$ and $(x_i^2)_{i=0}^{n}$ are depicted in Figure~\ref{fig:caval2}. Cavalieri's principle tells us that the area of integration strip $i$ is equal to $f(x_i^2)\Delta x_i^1$. Equation~\eqref{eq:cav_sum}, therefore, approximates the area of $R$. In the limit equation~\eqref{eq:cav_sum} approaches 
the Cavalieri integral (see Definition~\ref{def:cav_integral}):
\begin{equation}
\label{eq:caval1}
\int_{a(y)}^{b(y)}f(x)\, dx = \lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
If we evaluate the above integral we obtain the area of $R$. It is, however, quite hard to evaluate the above integral in its present form. Fortunately, it is easy to convert a Cavalieri integral into an equivalent Riemann or Riemann-Stieltjes 
integral by using the functions $h$ and $g$ (see equation~\eqref{eq:g_def} and equation~\eqref{eq:h_def}). Expressed mathematically (see Theorem~\ref{t:conv}):
\begin{equation}
\label{eq:main_cav}
\int_{a(y)}^{b(y)}f(x)\,dx =\int_a^b f \circ h (x)\, dx = \int_{a'}^{b'} f(x) dg(x),
\end{equation}
We can, therefore, evaluate equation~\eqref{eq:main_cav} in one of two ways:
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75,  
\end{equation}
or
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_{a'}^{b'} f \, dg(x) = \int_{\frac{1}{2}}^2x\, d2x = 3.75.  
\end{equation}
Note: $h(x) = \frac{1}{2}x$ and $g(x) = 2x$.\\

>>>>

\noindent
In this article we specifically focus on the geometric interpretation of fractional integrals. To be more specific, we present a novel geometric interpretation of the left-sided Riemann-Liouville fractional integral based at $a'$ (see Section~\ref{sec:ls}) \cite{laurent1884}. We accomplish this by extending the geometric framework proposed by Podlubny \cite{podlubny02}.
%As we mentioned earlier, Riemann and Liouville played an important role in the development of fractional calculus. Their work was crucial in defining what is now known as 
%the Riemann-Liouville fractional integral \cite{laurent1884}. At this point in time the reader should be made aware of the fact that there exists many other accepted fractional integral definitions in the literature.
%We, however, deem these definitions to be beyond the scope of the current article and as such we do not discuss them any further in this work. 
%The geometric interpretation that we present in this paper here is an extension of the geometric interpretation presented by Podlubny  
Podlubny showed that a Riemann-Liouville fractional integral can be converted into a Riemann-Stieltjes integral. The Riemann-Stieltjes integral so obtained can then be visualized using Bullock's geometric framework \cite{bullock88}. Moreover, Grobler showed that a Riemann-Stieltjes integral can be interpreted as the area which is obtained by summing together the area of an infinite number of non-rectangular infinitesimals, i.e. a Riemann-Stieltjes integral can be converted into a Cavalieri integral \cite{ackermann12,grobler19}. Combining these two ideas results in the following geometric interpretation: the Riemann-Liouville fractional integral $_{a'}I_t^{\alpha} f(t)$ (see Section~\ref{sec:ls}) can be interpreted as the area which is obtained by summing together the area of an infinite number of non-rectangular infinitesimals. The shape
of the aforementioned infinitesimals is determined by the order of integration $\alpha$ and the integration limit $t$ of the fractional integral being evaluated.\\

@@

\section{Riemann-Stieltjes integral}
\label{sec:rs}
\noindent
Let $f$ and $g$ denote real-valued functions defined on a closed interval $[a',b']$ of the real line. We shall suppose that both $f$ and $g$ are bounded on $[a',b']$. Recall that the Riemann-Stieltjes integral is defined as follows:
\begin{equation}
\label{eq:g_int2}
\int_{a'}^{b'} f(x) dg(x) =  \lim_{n \rightarrow \infty}\sum_{i=0}^{n-1} f(x_i)[g(x_{i+1})-g(x_{i})], 
\end{equation}
with $(x_i)_{i=0}^n$ being an arbitrary partition of $[a',b']$. The function $g$ is known as the integrator and if it is monotone increasing then it maps points in the interval $[a',b']$ to the interval $[a,b]$, with $a = g(a')$ and $b = g(b')$. The integral in equation~\eqref{eq:g_int2} is hard to evaluate in its present form. If $g$ is differentiable, however, then it becomes trivial to 
evaluate a Riemann-Stieltjes integral due to the following identity:
\begin{equation}
\label{eq:rs_identity}
\int_{a'}^{b'} f(x) dg(x) = \int_{a'}^{b'} f(x)g'(x)~dx.
\end{equation}

@@

\noindent
The definition of the Cavalieri integral is presented in Definition~\ref{def:cav_integral}. It is, however, our experience that Definition~\ref{def:cav_integral} is hard to assimilate if it is encountered by a reader for the first time. We, therefore, encourage readers who are not familiar with the Cavalieri integral to rather first read Section~\ref{sec:cav_integral_example} before continuing to read the current section. The reader who is unfamiliar with the Cavalieri integral should, however, refer back to the definitions presented in this section when things in Section~\ref{sec:cav_integral_example} become unclear. Moreover, the definitions presented below were taken from \cite{ackermann12}. Furthermore, none of the theorems in this section are proved, the proofs can be found in \cite{ackermann12,grobler19}.

@@

\noindent
The symbol $\circ$ denotes function composition. Let $a(y)$ be a translational function with respect to a function $f(x)$ on the interval $[a,b]$. Moreover, $b(y) = a(y) + (b-a)$. Furthermore, the functions $a(y)$ and $b(y)$ intersect $f(x)$ at $a'$ and $b'$ respectively. These assertions are assumed to be true throughout the paper and will not be repeated again.

\begin{definition}\label{def:h}
The mapping $h : [a, b] \rightarrow [a',b']$, which maps $x_i^1 \in [a, b]$ to $x_i^2 \in [a',b']$, is defined as
$h(x_i^1) :=$ $\{x_i^2 \in [a' ,b'] | a\circ f(x_i^2) + [x_i^1 - a] = x_i^2$ , $a = a(0)\}$
\end{definition}

\begin{definition}\label{def:g}
The mapping $g:[a', b'] \rightarrow [a, b]$, which maps $x_i^2 \in [a' , b']$ to $x_i^1\in [a, b]$,
is defined as $g(x_i^2) := x_i^2 - a \circ f (x_i^2) + a$.
\end{definition}

@@

\noindent
The notation $(\cdot)^{-1}$ denotes the functional inverse. The functions $h(x)$ and $g(x)$ are known as the transformation function and the inverse transformation function, respectively. Furthermore,
\begin{equation}
\label{eq:h_def}
h(x) = g^{-1}(x). 
\end{equation}
and
\begin{equation}
\label{eq:g_def}
g(x) = x - a\circ f(x) + a 
\end{equation}

\begin{definition}\label{def:cav_integral}
Define a partition $\mathcal{P}_1$ of $[a,b]$ to be a set of points $x_0^1, x_1^1,\cdots,x_n^1$, 
where $a = x_0^1 \leq x_1^1 \leq \cdots \leq x_n^1 = b$. For each partition 
$\mathcal{P}_1$ of $[a,b]$ write $\Delta x_k^1 = x_k^1-x_{k-1}^1$. 
%Since both boundaries of any integration strip are necessarily
%translations of the translational function $a(y)$, we can apply the transformation
%function $h$ to the partition $\mathcal{P}_1$. 
If the transformation function $h$ is strictly increasing,
the application of $h$ to the partition $\mathcal{P}_1$ induces a new partition $\mathcal{P}_2 = \{x_0^2, x_1^2,\cdots, x_n^2\}$,  
where $a' = x_0^2 \leq x_1^2 \leq \cdots \leq x_n^2 = b'$.
Otherwise, if $h$ is strictly decreasing, the application of $h$ induces a reversed partition $\mathcal{P}_2 = \{x_n^2, \cdots, x_1^2,x_0^2\}$. It
can be assumed that $h$ is strictly increasing, without any loss of generality.  Let 
$M_k = \sup \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, $m_k = \inf \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, and set 
\begin{equation}
\label{eq:c_up}
\mathcal{C}_U(\mathcal{P}_1,f,h) = \sum_{k=1}^n M_k \Delta x_k^1, 
\end{equation}
and
\begin{equation}
\label{eq:c_low}
\mathcal{C}_L(\mathcal{P}_1,f,h) = \sum_{k=1}^n m_k \Delta x_k^1. 
\end{equation}
The sums in equation~\eqref{eq:c_up} and equation~\eqref{eq:c_low} are respectively called the the upper and lower Cavalieri sums.
If there is a unique number $I$ that satisfies the inequality $\mathcal{C}_L(\mathcal{P}_1,f,h)\leq I \leq \mathcal{C}_U(\mathcal{P}_1,f,h)$ for all 
partitions $\mathcal{P}_1$ of $[a,b]$, then $I$ is called the Cavalieri integral of $f$ from $a(y)$ to $b(y)$ and is denoted by
\begin{equation}
\int_{a(y)}^{b(y)} f(x) dx.
\end{equation}
\end{definition}

@@

\noindent
Note that for all $K\in\mathbb{R}$ (see equation~\eqref{eq:rs_identity}):
\begin{equation}
\int_{a'}^{b'} f(x)\,d[\widetilde{g}(x)+K] = \int_{a'}^{b'}f(x)\,\widetilde{g}(x). 
\end{equation}
Furthermore, recall that (see equation~\ref{eq:g_def}):
\begin{equation}
\label{eq:g_definition}
g(x) = x - a\circ f(x) + a.
\end{equation}
Moreover, let $C$ be a real-valued constant that satisfies: $g(x) := \widetilde{g}(x)+C$.\\

@@

\section{Cavalieri Integral: Clarifying Example}
\label{sec:cav_integral_example}
\noindent
The example which we present below was taken from \cite{ackermann12,grobler19}. Let the region $R$ be bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$ and $b(y)=4-y$. This region is depicted in Figure~\ref{fig:caval2}. Let $a=1$, $b=4$, $a'=\frac{1}{2}$ and $b'=2$. Note that $a(y)$ is a translational function with respect to $f(x)$ on $[a,b]$. Moreover $a(y)$ and $b(y)$ intersect $f(x)$ at $a'$ and $b'$, respectively (see Definition~\ref{def:trans}).\\

@@

\noindent
Let us express this idea more formally. Let $(x_i^1)_{i=0}^{n}$ denote a partition on the $x$-axis, such that $a = x_0^1 < x_1^1 < \cdots < x_n^1 = b$, and $\Delta x_i^1 = x_{i+1}^1 - x_i^1$. Furthermore, consider the following lower Cavalieri sum (see Definition~\ref{def:cav_integral} and equation~\ref{eq:c_low}):
\begin{equation}
\label{eq:cav_sum}
\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
The partition points $(x_i^1)_{i=0}^{n}$ and $(x_i^2)_{i=0}^{n}$ are depicted in the left panel of Figure~\ref{fig:caval2} (if $n$ is set to 3). Cavalieri's principle tells us that the area of integration strip $i$ is equal to $f(x_i^2)\Delta x_i^1$. Equation~\eqref{eq:cav_sum}, therefore, approximates the area of $R$. In the limit equation~\eqref{eq:cav_sum} approaches 
the Cavalieri integral (see Definition~\ref{def:cav_integral}):
\begin{equation}
\label{eq:caval1}
\int_{a(y)}^{b(y)}f(x)\, dx = \lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
If we evaluate the above integral we obtain the area of $R$. It is, however, quite hard to evaluate the above integral in its present form. Fortunately, it is easy to convert a Cavalieri integral into an equivalent Riemann or Riemann-Stieltjes 
integral by using the functions $h$ and $g$ (see equation~\eqref{eq:g_def} and equation~\eqref{eq:h_def}). Expressed mathematically (see Theorem~\ref{t:conv}):
\begin{equation}
\label{eq:main_cav}
\int_{a(y)}^{b(y)}f(x)\,dx =\int_a^b f \circ h (x)\, dx = \int_{a'}^{b'} f(x) dg(x),
\end{equation}
We can, therefore, evaluate equation~\eqref{eq:main_cav} in one of two ways:
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75,  
\end{equation}
or
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_{a'}^{b'} f \, dg(x) = \int_{\frac{1}{2}}^2x\, d2x = 3.75.  
\end{equation}
Note: $h(x) = \frac{1}{2}x$ and $g(x) = 2x$.\\
